05-Exploring-avail_2021

### 2024-12-12 data-prep/01-processing-schools-parcels-census.qmd
changes to bib and analysis files after revision

Repeat for 2021 incomes:
```{r}
cen_data_2021_1 <- get_census(dataset='CA21', regions=list(CMA="35537"),
                           vectors=c("v_CA21_566", #Median after-tax income in 2020 among recipients ($),
                                     "v_CA21_966", #Median after-tax income of economic family in 2020 ($)
                                     "v_CA21_978", #Median after-tax income of 1-parent economic families in 2020 ($)
                                     "v_CA21_1043", #	Income; Low Income Measures; Prevalence of low income based on the Low-income measure, after tax (LIM-AT) (%); 0 to 17 years (%)
                                     "v_CA21_32", # population age 5-9
                                     "v_CA21_50", #population age 10-14
                                     "v_CA21_71"), #population age 15-19
                           level='DA', use_cache = FALSE, geo_format = 'sf')

@@

- Glen Echo, Glen Brae, Sir Isaac Brock, and Elizabeth Bagshaw elementary schools closed -- and "Viola Desmond Elementary School" opened in September 2021 on the site of Glen Brae. Let's make Glen Brae as "Removed", and add new school named "Viola Desmond" as SchoolID 156.202021, ELEM 0, urban.dist 7948.096, footprint 3,902 m²
(https://www.hwdsb.on.ca/sirwilfridlaurier/2021/05/19/viola-desmond-elementary-school-update/)

@@

Let's join add new columns, "Year_2" and "Status_2" and "footprint2021" corresponding to what happened between 2016-2021 for these 123 common schools.
```{r}
test_ALL_SCHOOLS_ELEM <- ALL_SCHOOLS_Elem |> mutate(Year_2 = ifelse(SchoolID %in% SCHOOLS_2024_and_2016$SchoolID, "2016 and 2021", NA),
                           Status_2 = ifelse(!is.na(Year_2), "NoChange", NA),
                           Status_2 = ifelse(SchoolID == "156", "Removed", Status_2),
                           footprint2021 = ifelse(SchoolID =="156", 3902, footprint2016))

test <- test_ALL_SCHOOLS_ELEM |> filter (SchoolID == "156") |> #takes the geometry and urban.dist of Glen Brae
  mutate(SchoolID = "156.2021",
         Name = "Viola Desmond",
         ELEM = "ELEM",
         MID = "MID",
         Year_2 = "2021",
         Status_2 = "New",
         footprint2021 =3902)

@@

Next assign 2016 DA IDs to the centroids of 2011 parcels:
```{r}
RED_LU_centroids_2011 <- sp::over(as_Spatial(st_centroid(RED_LU_2011)),as_Spatial(cen_data_2011_1 |> dplyr::select(GeoUID)))

@@

RED_LU_centroids_2011 <- cbind(RED_LU_centroids_2011_w2016ID,
                              RED_LU_centroids_2011,
                              st_centroid(RED_LU_2011))
```

Next assign 2016 DA IDs to the centroids of 2021 parcels:
```{r}
RED_LU_centroids_2021 <- sp::over(as_Spatial(st_centroid(RED_LU_2020)),
                                  as_Spatial(cen_data_2021_1 |> dplyr::select(GeoUID) |> st_sf()))

@@

RED_LU_centroids_2021 <- cbind(RED_LU_centroids_2021_w2016ID,
                              RED_LU_centroids_2021,
                              st_centroid(RED_LU_2020))
```

@@

Now add the TAZ IDs column to the residential parcel id for 2011, 2016, and 2021
```{r}
cols_TAZIDs <-sp::over(as_Spatial(st_centroid(RED_LU_2011)),
                              as_Spatial(ggh_taz |> select(GTA06) |> rename("GTA06_Orig" = "GTA06")))
RED_LU_centroids_2011<- cbind(RED_LU_centroids_2011,cols_TAZIDs)

cols_TAZIDs <- sp::over(as_Spatial(st_centroid(RED_LU_2016)),
                              as_Spatial(ggh_taz |> select(GTA06) |> rename("GTA06_Orig" = "GTA06")))

@@

Merge the the motor/non-motorized objects together:
```{r}
od_trips_perc_2011 <- od_trips_2011_motor %>% left_join(od_trips_2011_nonmotor, by=c("Origin", "Destination")) %>%
  group_by(Origin, Destination) |>
  summarize(TAZOrig_car_trips= sum(Trips.x, na.rm=TRUE),
            TAZOrig_walk_trips = sum(Trips.y, na.rm=TRUE)) |>
  mutate(TAZOrig_car_trip_perc = TAZOrig_car_trips/(TAZOrig_car_trips + TAZOrig_walk_trips))

od_trips_perc_2016 <- od_trips_2016_motor %>% left_join(od_trips_2016_nonmotor, by=c("Origin", "Destination")) %>%
  group_by(Origin,Destination) |>
  summarize(TAZOrig_car_trips = sum(Trips.x, na.rm=TRUE),
            TAZOrig_walk_trips = sum(Trips.y, na.rm=TRUE)) |>
  mutate(TAZOrig_car_trip_perc = TAZOrig_car_trips/(TAZOrig_car_trips + TAZOrig_walk_trips))
```

@@

>>>>

Repeat for 2021 incomes:
```{r}
cen_data_2021_1 <- get_census(dataset='CA21', regions=list(CMA="35537"),
                           vectors=c(
                                    "v_CA21_566", #Median after-tax income in 2020 among recipients ($),
                                     "v_CA21_966", #Median after-tax income of economic family in 2020 ($)
                                     "v_CA21_978", #Median after-tax income of 1-parent economic families in 2020 ($)
                                     "v_CA21_1043", #	Income; Low Income Measures; Prevalence of low income based on the Low-income measure, after tax (LIM-AT) (%); 0 to 17 years (%)
                                     "v_CA21_32", # population age 5-9
                                     "v_CA21_50", #population age 10-14
                                     "v_CA21_71"), #population age 15-19
                           level='DA', use_cache = FALSE, geo_format = 'sf')

@@

- Glen Echo, Glen Brae, Sir Isaac Brock, and Elizabeth Bagshaw elementary schools closed -- and "Viola Desmond Elementary School" opened in September 2021 on the site of Glen Brae. Let's make Glen Brae "Removed", and add new school named "Viola Desmond" as SchoolID 156.202021, ELEM 0, urban.dist 7948.096, footprint 3,902 m²
(https://www.hwdsb.on.ca/sirwilfridlaurier/2021/05/19/viola-desmond-elementary-school-update/)

@@

Let's join add new columns, "Year_2" and "Status_2" and "footprint2021" corresponding to what happened between 2016-2021 for these 123 common schools.
```{r}
test_ALL_SCHOOLS_ELEM <- ALL_SCHOOLS_Elem |>
  mutate(Year_2 = ifelse(SchoolID %in% SCHOOLS_2024_and_2016$SchoolID, "2016 and 2021", NA),
                           Status_2 = ifelse(!is.na(Year_2), "NoChange", NA),

                           Year_2 = ifelse(SchoolID == "156", "2016", Year_2),
                           Status_2 = ifelse(SchoolID == "156", "Removed", Status_2),
         footprint2021 = NA)

test <- test_ALL_SCHOOLS_ELEM |> filter (SchoolID == "156") |> #takes the geometry and urban.dist of Glen Brae
  mutate(SchoolID = "156.2021",
         Name = "Viola Desmond",
         ELEM = "ELEM",
         MID = "MID",
         Year_2 = "2021",
         Status_2 = "New",
         urban.dist = 7948.0958,
         footprint2021 =3902)

@@

Next assign 2016 DA IDs and 2021 DA IDs to the centroids of 2011 parcels:
```{r}
RED_LU_centroids_2011 <- sp::over(as_Spatial(st_centroid(RED_LU_2011)),as_Spatial(cen_data_2011_1 |> dplyr::select(GeoUID)))

@@

RED_LU_centroids_2011_w2021ID <- sp::over(as_Spatial(st_centroid(RED_LU_2011)),as_Spatial(cen_data_2021_1 |> dplyr::select(GeoUID) |> rename("GeoUID_2021" = "GeoUID")))

RED_LU_centroids_2011 <- cbind(RED_LU_centroids_2011_w2016ID,
                               RED_LU_centroids_2011_w2021ID,
                              RED_LU_centroids_2011,
                              st_centroid(RED_LU_2011))
```

And assign 2016 DA and 2011 IDs to the centroids of 2021 parcels:
```{r}
RED_LU_centroids_2021 <- sp::over(as_Spatial(st_centroid(RED_LU_2020)),
                                  as_Spatial(cen_data_2021_1 |> dplyr::select(GeoUID) |> st_sf()))

@@

RED_LU_centroids_2021_w2011ID <- sp::over(as_Spatial(st_centroid(RED_LU_2020)),
                                          as_Spatial(cen_data_2011_1 |> dplyr::select(GeoUID) |> rename("GeoUID_2011" = "GeoUID")))

RED_LU_centroids_2021 <- cbind(RED_LU_centroids_2021_w2016ID,
                               RED_LU_centroids_2021_w2011ID,
                              RED_LU_centroids_2021,
                              st_centroid(RED_LU_2020))
```

For good measure, do the same for the 2016 centroids:
```{r}
RED_LU_centroids_2016 <- sp::over(as_Spatial(st_centroid(RED_LU_2016)),
                                  as_Spatial(cen_data_2016_1 |> dplyr::select(GeoUID)))

RED_LU_centroids_2016_w2011ID <- sp::over(as_Spatial(st_centroid(RED_LU_2016)),as_Spatial(cen_data_2011_1 |> dplyr::select(GeoUID) |> rename("GeoUID_2011" = "GeoUID")))

RED_LU_centroids_2016_w2021ID <- sp::over(as_Spatial(st_centroid(RED_LU_2016)),as_Spatial(cen_data_2021_1 |> dplyr::select(GeoUID) |> rename("GeoUID_2021" = "GeoUID")))

RED_LU_centroids_2016 <- cbind(RED_LU_centroids_2016_w2011ID,
                               RED_LU_centroids_2016_w2021ID,
                              RED_LU_centroids_2016,
                              st_centroid(RED_LU_2016))
```

@@

Now add the TAZ IDs column to the residential parcel id for 2011, 2016, and 2021
```{r}
cols_TAZIDs <-sp::over(as_Spatial(st_centroid(RED_LU_2011)),
                              as_Spatial(ggh_taz |> select(GTA06) |> rename("GTA06_Orig" = "GTA06")))

RED_LU_centroids_2011<- cbind(RED_LU_centroids_2011,cols_TAZIDs)

cols_TAZIDs <-sp::over(as_Spatial(st_centroid(RED_LU_2016)),
                              as_Spatial(ggh_taz |> select(GTA06) |> rename("GTA06_Orig" = "GTA06")))

@@

Merge the the motor/non-motorized objects together:
```{r}
od_trips_perc_2011 <- od_trips_2011_motor %>% full_join(od_trips_2011_nonmotor, by=c("Origin", "Destination")) %>%
  group_by(Origin, Destination) |>
  summarize(TAZOrig_car_trips= sum(Trips.x, na.rm=TRUE),
            TAZOrig_walk_trips = sum(Trips.y, na.rm=TRUE)) |>
  mutate(TAZOrig_car_trip_perc = TAZOrig_car_trips/(TAZOrig_car_trips + TAZOrig_walk_trips))

od_trips_perc_2016 <- od_trips_2016_motor %>% full_join(od_trips_2016_nonmotor, by=c("Origin", "Destination"))%>%
  group_by(Origin,Destination) |>
  summarize(TAZOrig_car_trips = sum(Trips.x, na.rm=TRUE),
            TAZOrig_walk_trips = sum(Trips.y, na.rm=TRUE)) |>
  mutate(TAZOrig_car_trip_perc = TAZOrig_car_trips/(TAZOrig_car_trips + TAZOrig_walk_trips))
```

@@

```{r}
##TEEESSST -- here we make sure that the full_join worked correctly for calculating the % of trips per mode.. it works!
st_intersection(st_centroid(ham_taz) %>% st_transform(crs = 4326), 
                ham_community_bounds%>% st_transform(crs = 4326)) |> filter(COMMUNITY_ == "Dundas")

TEST <-c(5178, 5191, 5202, 5205, 5207, 5213, 5215, 5236, 5246, 5247, 5252)
```

```{r}
#see, 29 rows --> it comes from 25 rows and 7 rows (added together = 32 + collapsing 3 rows as those ODs both have walk and car trips = 29 unique ODs!)
od_trips_perc_2011 |> filter(Origin %in% TEST)
od_trips_2011_motor |> filter(Origin %in% TEST)
od_trips_2011_nonmotor |> filter(Origin %in% TEST)
```

### 2024-12-12 data-prep/04-Creating-avail_2011-16.Rmd
Added 2021 schools and census to the database. Started Manuscript changes.
changes to bib and analysis files after revision

---
title: "Testing-doc-2016run"
output: html_document
date: "2024-02-16"
---
# 2011 calcs

@@

```{r}
load(file="data-products/cen_data_2011_hamiltononly.Rdata")
cen_data_2011_hamiltononly <- cen_data_2011_1 |> filter(Region.Name == "Hamilton") #NOTE: GEOUID 35250242, 35250133, 35250457 do not have any child population
rm("cen_data_2011_1")
load(file="data-products/cen_data_2016_hamiltononly.Rdata")
cen_data_2016_hamiltononly <- cen_data_2016_1 |> filter(Region.Name == "Hamilton") #NOTE: GEOUID 35250133, 35250398, 35250427 do not have any child population
rm("cen_data_2016_1")

ham_city_bound <- st_read("data-inputs/Boundaries/City_Boundary.shp") |> st_transform(st_crs(cen_data_2011_hamiltononly))
ham_community_bounds <- st_read("data-inputs/Boundaries/Community_Boundaries.shp") |> st_transform(st_crs(cen_data_2011_hamiltononly))
ggh_taz <- st_read("data-inputs/TTS/tts06_83_region.shp") |> st_transform(st_crs(cen_data_2011_hamiltononly))
```

```{r}
cen_data_2011_hamiltononly <- 
  cen_data_2011_hamiltononly %>% dplyr::select( #removing the variables we don't need for this analysis
  -c("Shape.Area", "Quality.Flags", "Type", "CD_UID", "NHS.Non.Return.Rate", "CSD_UID", "CT_UID", "CT_UID","NHS.Non.Return.Rate.1", "v_CA11N_2564..Median.after.tax.household.income..", "v_CA11F_17..15.to.19.years")) %>%
  rename( #renaming the variables
    "DA_households" = "Households",
    "DA_dwellings" = "Dwellings",
    "DA_tot_pop" = "Population",
    "DA_med_AT_family_income" = "v_CA11N_2458..Median.after.tax.family.income..",
    "DA_med_AT_lone-family_income" = "v_CA11N_2476..Median.after.tax.family.income..",
    "DA_prev_LIMAT_under18" = "v_CA11N_2609..Less.than.18.years..",
    "DA_pop_5to9" = "v_CA11F_11..5.to.9.years",
    "DA_pop_10to14" = "v_CA11F_14..10.to.14.years") %>%
  mutate(DA_pop_5to14 = DA_pop_5to9+(DA_pop_10to14)) |> #*0.9)) %>%
  dplyr::select(-c("DA_pop_5to9","DA_pop_10to14"))

cen_data_2016_hamiltononly <- 
  cen_data_2016_hamiltononly %>% dplyr::select( #removing the variables we don't need for this analysis
  -c("Shape.Area",  "Type", "CD_UID", "CSD_UID", "CT_UID", "CT_UID", "v_CA16_2398..Median.after.tax.income.of.households.in.2015....", "v_CA16_64..15.to.19.years")) %>%
  rename( #renaming the variables
    "DA_households" = "Households",
    "DA_dwellings" = "Dwellings",
    "DA_tot_pop" = "Population",
    "DA_med_AT_family_income" = "v_CA16_2448..Median.after.tax.income.of.economic.families.in.2015....",
         "DA_med_AT_lone-family_income" = "v_CA16_2460..Median.after.tax.income.of.lone.parent.economic.families.in.2015....",
         "DA_prev_LIMAT_under18" = "v_CA16_2543..0.to.17.years....",
         "DA_pop_5to9" = "v_CA16_25..5.to.9.years",
         "DA_pop_10to14" = "v_CA16_43..10.to.14.years") %>% mutate(DA_pop_5to14 = DA_pop_5to9+(DA_pop_10to14)) |> #*0.9)) %>%
         dplyr::select(-c("DA_pop_5to9","DA_pop_10to14"))
```

```{r}
library(hamONdests)
data(Schools_2011_2016)
ALL_SCHOOLS_Elem <- Schools_2011_2016 %>% filter(Level == "Elementary")
rm("Schools_2011_2016")
#data(School_Catchments_2011_2016)
load(file="data-products/TTS_gamma2011_motor.Rdata")
load(file="data-products/TTS_gamma2016_motor.Rdata")
load(file="data-products/TTS_exp2011_nonmotor.Rdata")
load(file="data-products/TTS_exp2016_nonmotor.Rdata")
```

```{r}
#add GTA06 IDs to schools - "GTA06_Dest"}
cols_TAZIDs <- sp::over(as_Spatial(ALL_SCHOOLS_Elem |> st_transform(st_crs(ggh_taz))),
                              as_Spatial(ggh_taz |> dplyr::select(GTA06) |> rename("GTA06_Dest" = "GTA06")))

ALL_SCHOOLS_Elem <- cbind(ALL_SCHOOLS_Elem,cols_TAZIDs)
rm(cols_TAZIDs)
```

```{r}
load(file="data-products/OD_tt_car_2011_1.Rdata")
load(file="data-products/OD_tt_car_2011_2.Rdata")
load(file="data-products/OD_tt_car_2011_3.Rdata")
load(file="data-products/OD_tt_car_2011_4.Rdata")

@@

```{r}
#create OD matrix -- each parcel to each school, and then drop all OD pairs that don't have a travel time ()
OD_2011 <- RED_LU_centroids_2011 %>% 
  cross_join(ALL_SCHOOLS_Elem %>% st_drop_geometry() |> filter(Year != 2016) %>% dplyr::select(c("SchoolID", "OTGC2011", "GTA06_Dest")))

@@

OD_2011 <- rbind(test,OD_2011) 
rm(test, ttm_car_2011, ttm_walk_2011)
```
```{r}
#reduce dimensions
OD_2011 <- OD_2011 %>% dplyr::select(-c("LUC1_DESC", "LUC2_DESC", "res_units", "DA_dwellings", "CMA_UID", "Region.Name", "geometry","PROP_AREA", "Area..sq.km."))

@@

# 2016 calcs
```{r}
rm(list=ls())
```

```{r load-packages, include=FALSE, cache=FALSE}
library(dplyr)
library(sf)
library(ggplot2)
library(tidyverse)
library(units)
library(patchwork)
library(ggspatial)
options(scipen = 999)
```

```{r sp_avail-function,include=FALSE}
#defining the spatial availability function, this prints all the intermediates
sp_avail_detailed <- function(x, o_id, d_id, pop, opp, pop_type, r, f, alpha = 1){

```{r}
load(file="data-products/cen_data_2016_hamiltononly.Rdata")
cen_data_2016_hamiltononly <- cen_data_2016_1 |> filter(Region.Name == "Hamilton") #NOTE: GEOUID 35250133, 35250398, 35250427 do not have any child population
rm("cen_data_2016_1")

ham_city_bound <- st_read("data-inputs/Boundaries/City_Boundary.shp") |> st_transform(st_crs(cen_data_2016_hamiltononly))
ham_community_bounds <- st_read("data-inputs/Boundaries/Community_Boundaries.shp") |> st_transform(st_crs(cen_data_2016_hamiltononly))
ggh_taz <- st_read("data-inputs/TTS/tts06_83_region.shp") 
# |> st_transform(st_crs(cen_data_2011_hamiltononly))
```

```{r}
cen_data_2016_hamiltononly <- 
  cen_data_2016_hamiltononly %>% dplyr::select( #removing the variables we don't need for this analysis
  -c("Shape.Area",  "Type", "CD_UID", "CSD_UID", "CT_UID", "CT_UID", "v_CA16_2398..Median.after.tax.income.of.households.in.2015....", "v_CA16_64..15.to.19.years")) %>%
  rename( #renaming the variables
    "DA_households" = "Households",
    "DA_dwellings" = "Dwellings",
    "DA_tot_pop" = "Population",
    "DA_med_AT_family_income" = "v_CA16_2448..Median.after.tax.income.of.economic.families.in.2015....",
         "DA_med_AT_lone-family_income" = "v_CA16_2460..Median.after.tax.income.of.lone.parent.economic.families.in.2015....",
         "DA_prev_LIMAT_under18" = "v_CA16_2543..0.to.17.years....",
         "DA_pop_5to9" = "v_CA16_25..5.to.9.years",
         "DA_pop_10to14" = "v_CA16_43..10.to.14.years") %>% mutate(DA_pop_5to14 = DA_pop_5to9+(DA_pop_10to14)) |> #*0.9)) %>%
         dplyr::select(-c("DA_pop_5to9","DA_pop_10to14"))
```

```{r, include=FALSE}
library(hamONdests)
data(Schools_2011_2016)
ALL_SCHOOLS_Elem <- Schools_2011_2016 %>% filter(Level == "Elementary")
rm("Schools_2011_2016")
#data(School_Catchments_2011_2016)
load(file="data-products/TTS_gamma2016_motor.Rdata")
load(file="data-products/TTS_exp2016_nonmotor.Rdata")
```

```{r, include=FALSE}
#add GTA06 IDs to schools - "GTA06_Dest"}
cols_TAZIDs <- sp::over(as_Spatial(ALL_SCHOOLS_Elem |> st_transform(st_crs(ggh_taz))),
                              as_Spatial(ggh_taz |> dplyr::select(GTA06) |> rename("GTA06_Dest" = "GTA06")))
ALL_SCHOOLS_Elem <- cbind(ALL_SCHOOLS_Elem,cols_TAZIDs)
rm(cols_TAZIDs)
```

@@

```{r}
#create OD matrix -- each parcel to each school, and then drop all OD pairs that don't have a travel time ()
OD_2016 <- RED_LU_centroids_2016 %>% 
  cross_join(ALL_SCHOOLS_Elem %>% st_drop_geometry() |> filter(Year != 2011) %>% dplyr::select(c("SchoolID", "OTGC2016", "GTA06_Dest")))

@@

OD_2016 <- rbind(test,OD_2016) 
rm(test, ttm_car_2016, ttm_walk_2016)
```
```{r}
#reduce dimensions
OD_2016 <- OD_2016 %>% dplyr::select(-c("LUC1_DESC", "LUC2_DESC", "res_units", "DA_dwellings", "CMA_UID", "Region.Name", "geometry","PROP_AREA", "Area..sq.km."))

@@

```{r}
#checks!, these should be the total number of kids (^the code above) -- 58265.
OD_2016$kid_pop %>% sum()
OD_2016$kid_popadj %>% sum()
OD_2016 |> group_by(GeoUID) |> summarize(DA_pop_5to14 = mean(DA_pop_5to14)) |> dplyr::select(DA_pop_5to14) |> sum()
```

>>>>

---
title: "Testing-doc-2016run"
output: html_document
date: "2024-02-16"
---

@@

All schools, including 2024 schools. Supercedes the hamonDESTs package:
```{r}
load(file="data-products/SCHOOLS.Rdata") #for this paper, the schools under analysis
```

Load census
```{r}
load(file="data-products/cen_data_2011_hamiltononly.Rdata")
cen_data_2011_hamiltononly <- cen_data_2011_1 |> filter(`Region Name` == "Hamilton") #NOTE: GEOUID 35250242, 35250133, 35250457 do not have any child population
rm("cen_data_2011_1")
load(file="data-products/cen_data_2016_hamiltononly.Rdata")
cen_data_2016_hamiltononly <- cen_data_2016_1 |> filter(`Region Name` == "Hamilton") #NOTE: GEOUID 35250133, 35250398, 35250427 do not have any child population
rm("cen_data_2016_1")
load(file="data-products/cen_data_2021_hamiltononly.Rdata")
cen_data_2021_hamiltononly <- cen_data_2021_1 
rm("cen_data_2021_1")

ham_city_bound <- st_read("data-inputs/Boundaries/City_Boundary.shp") |> st_transform(st_crs(cen_data_2011_hamiltononly))
ham_community_bounds <- st_read("data-inputs/Boundaries/Community_Boundaries.shp") |> st_transform(st_crs(cen_data_2011_hamiltononly))
ggh_taz <- st_read("data-inputs/TTS/tts06_83_region.shp") |> st_transform(st_crs(cen_data_2011_hamiltononly))
```
```{r}
#buffering to clean up the edges of these objects -> there wre issues using st_join. 
sf_use_s2(FALSE)
ggh_taz <- ggh_taz |> st_make_valid()
ggh_taz <- st_buffer(ggh_taz, dist=0)
ham_taz <- ggh_taz %>% dplyr::filter(REGION==6) #Filter out everything other than Planning District 6 = Hamilton
```

```{r}
cen_data_2011_hamiltononly <- 
  cen_data_2011_hamiltononly %>% dplyr::select( #removing the variables we don't need for this analysis
  -c("Shape Area", "Quality Flags", "Type", "CD_UID", "NHS Non-Return Rate", "CSD_UID", "CT_UID", "CT_UID","NHS Non Return Rate", "v_CA11N_2564: Median after-tax household income $", "v_CA11F_17: 15 to 19 years")) %>%
  rename( #renaming the variables
    "DA_households" = "Households",
    "DA_dwellings" = "Dwellings",
    "DA_tot_pop" = "Population",
    "DA_med_AT_family_income" = "v_CA11N_2458: Median after-tax family income $",
    "DA_med_AT_lone-family_income" = "v_CA11N_2476: Median after-tax family income $",
    "DA_prev_LIMAT_under18" = "v_CA11N_2609: Less than 18 years %",
    "DA_pop_5to9" = "v_CA11F_11: 5 to 9 years",
    "DA_pop_10to14" = "v_CA11F_14: 10 to 14 years") %>%
  mutate(DA_pop_5to14 = DA_pop_5to9+(DA_pop_10to14)) |> #*0.9)) %>%
  dplyr::select(-c("DA_pop_5to9","DA_pop_10to14"))
```

```{r}
cen_data_2016_hamiltononly <- 
  cen_data_2016_hamiltononly %>% dplyr::select( #removing the variables we don't need for this analysis
  -c("Shape Area",  "Type", "CD_UID", "CSD_UID", "CT_UID", "CT_UID", "v_CA16_2398: Median after-tax income of households in 2015 ($)",  "v_CA16_64: 15 to 19 years"  )) %>%
  rename( #renaming the variables
    "DA_households" = "Households",
    "DA_dwellings" = "Dwellings",
    "DA_tot_pop" = "Population",
    "DA_med_AT_family_income" = "v_CA16_2448: Median after-tax income of economic families in 2015 ($)",
    "DA_med_AT_lone-family_income" = "v_CA16_2460: Median after-tax income of lone-parent economic families in 2015 ($)",
         "DA_prev_LIMAT_under18" = "v_CA16_2543: 0 to 17 years (%)",
         "DA_pop_5to9" = "v_CA16_25: 5 to 9 years",
         "DA_pop_10to14" = "v_CA16_43: 10 to 14 years") %>% mutate(DA_pop_5to14 = DA_pop_5to9+(DA_pop_10to14)) |> #*0.9)) %>%
         dplyr::select(-c("DA_pop_5to9","DA_pop_10to14"))
```

```{r}
cen_data_2021_hamiltononly <- 
  cen_data_2021_hamiltononly %>% dplyr::select( #removing the variables we don't need for this analysis
  -c("Shape Area",  "Type", "CD_UID", "CSD_UID", "CT_UID", "CT_UID", "v_CA21_566: Median after-tax income in 2020 among recipients ($)" ,"v_CA21_71: 15 to 19 years")) %>%
  rename( #renaming the variables
    "DA_households" = "Households",
    "DA_dwellings" = "Dwellings",
    "DA_tot_pop" = "Population",
    "DA_med_AT_family_income" = "v_CA21_966: Median after-tax income of economic family in 2020 ($)",
    "DA_med_AT_lone-family_income" = "v_CA21_978: Median after-tax income of one-parent economic families in 2020 ($)",
    "DA_prev_LIMAT_under18" =  "v_CA21_1043: 0 to 17 years",
    "DA_pop_5to9" = "v_CA21_32: 5 to 9 years",
    "DA_pop_10to14" = "v_CA21_50: 10 to 14 years") %>%
    mutate(DA_pop_5to14 = DA_pop_5to9+(DA_pop_10to14)) |> #*0.9)) %>%
         dplyr::select(-c("DA_pop_5to9","DA_pop_10to14"))
```

load TTS travel impedance functions, and format SCHOOLS to refer to the TTS IDS (add GTA06 IDs to schools - "GTA06_Dest"):
```{r}
load(file="data-products/TTS_gamma2011_motor.Rdata")
load(file="data-products/TTS_exp2011_nonmotor.Rdata")
load(file="data-products/TTS_gamma2016_motor.Rdata")
load(file="data-products/TTS_exp2016_nonmotor.Rdata")

#formatting
cols_TAZIDs <- sp::over(as_Spatial(SCHOOLS |> st_transform(st_crs(ggh_taz))),
                              as_Spatial(ggh_taz |> dplyr::select(GTA06) |> rename("GTA06_Dest" = "GTA06")))

SCHOOLS <- cbind(SCHOOLS,cols_TAZIDs)
rm(cols_TAZIDs)
```

# 2011 calcs
```{r}
load(file="data-products/OD_tt_car_2011_1.Rdata")
load(file="data-products/OD_tt_car_2011_2.Rdata")
load(file="data-products/OD_tt_car_2011_3.Rdata")
load(file="data-products/OD_tt_car_2011_4.Rdata")

@@

```{r}
#create OD matrix -- each parcel to each school, and then drop all OD pairs that don't have a travel time ()
OD_2011 <- RED_LU_centroids_2011 %>% 
  cross_join(SCHOOLS %>% st_drop_geometry() |> filter(Year != 2016) %>% dplyr::select(c("SchoolID", "OTGC2011", "GTA06_Dest")))

@@

OD_2011 <- rbind(test,OD_2011) 
rm(test, ttm_car_2011, ttm_walk_2011)
```
```{r}
#reduce dimensions
OD_2011 <- OD_2011 %>% dplyr::select(-c("LUC1_DESC", "LUC2_DESC", "DA_dwellings", "CMA_UID", "Region Name", "geometry","Area (sq km)"))

@@

# 2016 calcs
```{r}
rm(avail_2011,ttm_car_2011, ttm_walk_2011, od_trips_perc_2011, od_trips_perc_comm_2011,RED_LU_centroids_2011)
```

@@

```{r}
#create OD matrix -- each parcel to each school, and then drop all OD pairs that don't have a travel time ()
OD_2016 <- RED_LU_centroids_2016 %>% 
  cross_join(SCHOOLS %>% st_drop_geometry() |> filter(Year != 2011) %>% dplyr::select(c("SchoolID", "OTGC2016", "GTA06_Dest")))

@@

OD_2016 <- rbind(test,OD_2016) 
rm(test, ttm_car_2016, ttm_walk_2016)
```
```{r}
#reduce dimensions
OD_2016 <- OD_2016 %>% dplyr::select(-c("LUC1_DESC", "LUC2_DESC", "res_units", "DA_dwellings", "CMA_UID", "Region Name", "geometry","PROP_AREA", "Area (sq km)"))

@@

```{r}
#checks!, these should be the total number of kids (^the code above) -- 58865
OD_2016$kid_pop %>% sum()
OD_2016$kid_popadj %>% sum()
OD_2016 |> group_by(GeoUID) |> summarize(DA_pop_5to14 = mean(DA_pop_5to14)) |> dplyr::select(DA_pop_5to14) |> sum()
```

### 2024-12-12 data-prep/05-Exploring-avail_2021.qmd
changes to bib and analysis files after revision

---
title: "05-Exploring-avail_2021"
format: html
---
library:
```{r}
#install.packages("sf")
library(dplyr)
library(sf)
# if(!require(remotes)){
#     install.packages("remotes")
#     library(remotes)
# }
#remotes::install_github("soukhova/HamONdests")
library(hamONdests)
```

@@

#drop all parcel points that do not have a GeoUID (i.e., a DA identifier -- this means they fall outside of Hamilton). It's 10 for 2011,  11 for 2016, and 11 for 2021
RED_LU_centroids_2021 <- RED_LU_centroids_2021 %>% drop_na(GeoUID) %>%
  group_by(ID) %>%
  dplyr::select(c("ID","GeoUID_2016","GeoUID","GTA06_Orig","WARD","COMMUNITY","URBAN_RURA","LUC1","LUC2","LUC1_DESCR","TYPE","geometry"))
```

@@

```{r}
#calculate spatial availability. 
avail_2021 <- sp_avail_detailed(OD_2021, ID, SchoolID, kid_popadj, OTGC2021, mode, r, f, alpha = 1)
save(avail_2021, file="data-products/avail_2021.RData")
save(OD_2021, file="data-products/OD_2021.RData")
```

```{r}
#checks, yes!
OD_2021$OTGC2021 |> unique() |> sum()
avail_2021$V_ij |> sum(na.rm=T)
```

@@

>>>>

---
title: "05-Exploring-avail_2021"
format: html
---
library:
```{r}
#install.packages("sf")
library(dplyr)
library(sf)
library(tidyr)
# if(!require(remotes)){
#     install.packages("remotes")
#     library(remotes)
# }
# remotes::install_github("soukhova/TTS2016R")
library(hamONdests)
```

@@

#drop all parcel points that do not have a GeoUID (i.e., a DA identifier -- this means they fall outside of Hamilton). It's 10 for 2011,  11 for 2016, and 11 for 2021
RED_LU_centroids_2021 <- RED_LU_centroids_2021 %>% drop_na(GeoUID) %>%
  group_by(ID) %>%
  dplyr::select(c("ID","GeoUID", "GeoUID_2011","GeoUID_2016",
                  "GTA06_Orig","WARD","COMMUNITY","URBAN_RURA","LUC1","LUC2","LUC1_DESCR","TYPE","geometry"))
```

@@

```{r}
#calculate spatial availability. 
avail_2021 <- sp_avail_detailed(OD_2021, ID, SchoolID, kid_popadj, OTGC2021, mode, r, f, alpha = 1)
```

```{r}
save(avail_2021, file="data-products/avail_2021.RData")
save(OD_2021, file="data-products/OD_2021.RData")
```

```{r}
#checks, yes!
OD_2021$OTGC2021 |> unique() |> sum()
avail_2021$V_ij |> sum(na.rm=T)
#should be  57696.08!
```

@@

Also interesting! The parcels have "urban vs. rural" in 2021... let's summarise this per DA and save in data_products.
```{r}
load(file="data-inputs/Parcels/RED_LU_centroids_2021.RData") #parcels

DA_2021_URBvsRUR <- RED_LU_centroids_2021 |> st_drop_geometry() |>
  mutate(URB = ifelse(URBAN_RURA == "U", 1, 0),
         SUB = ifelse(URBAN_RURA == "S", 1, 0),
         RUR = ifelse(URBAN_RURA == "R", 1, 0)) |>
  group_by(GeoUID) |> summarise(count_urb= sum(URB),
                                count_sub = sum(SUB),
                                count_rur = sum(RUR))

DA_2021_URBvsRUR <- DA_2021_URBvsRUR |>
  transmute(GeoUID,
            perc_urb = count_urb/rowSums(DA_2021_URBvsRUR[2:4]))

save(DA_2021_URBvsRUR, file="data-products/DA_2021_URBvsRUR.RData")
```
